---
title: "exploratory_analysis"
author: 
date: "2022-12-01"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)
library(ggplot2)
library(patchwork)
library(dplyr)
library(readxl)
library(tidycensus)
library(rvest)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
```

```{r load datasets}
menustat_df = read_csv("data/menustat_combined.csv")
rest_locations = read_csv("data/restaurant_location.csv") 
```

# Box plot of nutritional content by restaurant type across combined years (2008-2020)

```{r}
menustat_df %>% 
  group_by(rest_type) %>% ##why am i getting an NA for restaurant type?
  drop_na(rest_type) %>% 
  ggplot(aes(x = rest_type, y = calories)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Calories Per Food Item",
    title = "Distribution of Calories in Food Items Served by Restaurant Type") 

menustat_df %>% 
  distinct(rest_type)

```
## what items are over 5000 calories each?

```{r}
cal_dense = 
  menustat_df %>% 
  mutate(
    as.numeric(calories)
  ) %>% 
  filter(
    calories > 5000
  )

fast_cal_dense = 
  cal_dense %>% 
  filter(
    rest_type == "fast food"
  ) %>% 
  select(rest_type, item_name, calories)

fast_cal_dense_plot =
  fast_cal_dense %>% 
  ggplot(aes(x = item_name, y = calories)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 6)) +
  labs(
    x = "Item Name",
    y = "Calories per Item",
    title = "Food Items > 5500 calories in Fast Food Restaurants"
  ) 

fast_cal_dense_plot

sit_cal_dense = 
  cal_dense %>% 
  filter(
    rest_type == "sit-down"
  ) %>% 
  select(rest_type, item_name, calories)
  
sit_cal_dense_plot =
sit_cal_dense %>% 
  ggplot(aes(x = item_name, y = calories)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 6)) +
  labs(
    x = "Item Name",
    y = "Calories per Item",
    title = "Food Items > 5000 calories in Sit-Down Restaurants"
  ) 

sit_cal_dense_plot

(fast_cal_dense_plot + sit_cal_dense_plot)
  
cal_dense %>% 
  ggplot(aes(x = item_name, y = calories, color = rest_type)) +
  geom_point(alpha = .5) +
  theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 6)) +
  labs(
    x = "Item Name",
    y = "Calories per Item",
    title = "Scatterplot of Food Items > 5000 calories "
  ) 

## again i think these would read better if interactive -> look into making a flex dashboard that has more than 3 plots...
## dropping out the items that are in the top 5% because these items are all meant to share and are not individual portion sizes having this many calories 

final_cal_df = 
  menustat_df %>% 
  group_by(rest_type) %>% 
  filter(quantile(calories, 0.95) > calories)

final_cal_df %>% 
  ggplot(aes(x = rest_type, y = calories)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Calories Per Food Item",
    title = "Distribution of Calories in Food Items Served by Restaurant Type")
```

## let's look at if distribution of calories changed over the years (2008, 2012, 2016, 2020)

```{r}
cal_2008 =
  final_cal_df %>% 
  filter(
    year == "2008"
  )

cal_08_plot =
  cal_2008 %>% 
  ggplot(aes(x = rest_type, y = calories)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Calories Per Food Item",
    title = "Distribution of Calories in Food Items Served by Restaurant Type in 2008") 

cal_08_plot

cal_2012 =
  final_cal_df %>% 
  filter(
    year == "2012"
  )

cal_12_plot =
  cal_2012 %>% 
  ggplot(aes(x = rest_type, y = calories)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Calories Per Food Item",
    title = "Distribution of Calories in Food Items Served by Restaurant Type in 2012") 

cal_12_plot

cal_2016 =
  final_cal_df %>% 
  filter(
    year == "2016"
  )

cal_16_plot =
  cal_2016 %>% 
  ggplot(aes(x = rest_type, y = calories)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Calories Per Food Item",
    title = "Distribution of Calories in Food Items Served by Restaurant Type in 2016") 

cal_16_plot

cal_2020 =
  final_cal_df %>% 
  filter(
    year == "2020"
  )

cal_20_plot =
  cal_2020 %>% 
  ggplot(aes(x = rest_type, y = calories)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Calories Per Food Item",
    title = "Distribution of Calories in Food Items Served by Restaurant Type in 2020") 

cal_20_plot

(cal_08_plot +  cal_12_plot + cal_16_plot + cal_20_plot)

## use patchwork to put all the years side by side 
## i want this to be interactive so that people can hover and see what item is that many calories 
## add another link on your personal website to mess around with the flex dashboard 
```
In 2008 see that the max calorie item recorded was almost 3000 cals. 
In 2016 see the NAs come back....huh... 

## boxplot of changing distribution of fat content over the years 

```{r}
fat_2008 = 
  final_cal_df %>% 
  mutate(
    as.numeric(trans_fat)
  ) %>% 
  filter(
    year == "2008")

  
fat_08_plot =
  fat_2008 %>% 
  ggplot(aes(x = rest_type, y = trans_fat)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Trans Fat Per Food Item",
    title = "Distribution of Trans Fat in Food Items Served by Restaurant Type in 2008") 

fat_08_plot

fat_2012 = 
  final_cal_df %>% 
  mutate(
    as.numeric(trans_fat)
  ) %>% 
  filter(
    year == "2012")

  
fat_12_plot =
  fat_2012 %>% 
  ggplot(aes(x = rest_type, y = trans_fat)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Trans Fat Per Food Item",
    title = "Distribution of Trans Fat in Food Items Served by Restaurant Type in 2012") 

fat_12_plot


fat_2016 = 
  final_cal_df %>% 
  mutate(
    as.numeric(trans_fat)
  ) %>% 
  filter(
    year == "2016")

  
fat_16_plot =
  fat_2016 %>% 
  ggplot(aes(x = rest_type, y = trans_fat)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Trans Fat Per Food Item",
    title = "Distribution of Trans Fat in Food Items Served by Restaurant Type in 2016") 

fat_16_plot


fat_2020 = 
  final_cal_df %>% 
  mutate(
    as.numeric(trans_fat)
  ) %>% 
  filter(
    year == "2020")

  
fat_20_plot =
  fat_2020 %>% 
  ggplot(aes(x = rest_type, y = trans_fat)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Trans Fat Per Food Item",
    title = "Distribution of Trans Fat in Food Items Served by Restaurant Type in 2020") 

fat_20_plot


(fat_08_plot +  fat_12_plot + fat_16_plot + fat_20_plot)
```
### boxplot of protein content over the years by restaurant type 

```{r}
prot_2008 = 
  final_cal_df %>% 
  mutate(
    as.numeric(protein)
  ) %>% 
  filter(
    year == "2008")

  
prot_08_plot =
  prot_2008 %>% 
  ggplot(aes(x = rest_type, y = protein)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Protein Per Food Item",
    title = "Distribution of Protein in Food Items Served by Restaurant Type in 2008") 

prot_08_plot

prot_2012 = 
  final_cal_df %>% 
  mutate(
    as.numeric(protein)
  ) %>% 
  filter(
    year == "2012")

  
prot_12_plot =
  prot_2012 %>% 
  ggplot(aes(x = rest_type, y = protein)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Protein Per Food Item",
    title = "Distribution of Protein in Food Items Served by Restaurant Type in 2012") 

prot_12_plot


prot_2016 = 
  final_cal_df %>% 
  mutate(
    as.numeric(protein)
  ) %>% 
  filter(
    year == "2016")

  
prot_16_plot =
  prot_2016 %>% 
  ggplot(aes(x = rest_type, y = protein)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Protein Per Food Item",
    title = "Distribution of Protein in Food Items Served by Restaurant Type in 2016") 

prot_16_plot


prot_2020 = 
  final_cal_df %>% 
  mutate(
    as.numeric(protein)
  ) %>% 
  filter(
    year == "2020")

  
prot_20_plot =
  prot_2020 %>% 
  ggplot(aes(x = rest_type, y = protein)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Protein Per Food Item",
    title = "Distribution of Protein in Food Items Served by Restaurant Type in 2020") 

prot_20_plot


(prot_08_plot +  prot_12_plot + prot_16_plot + prot_20_plot)
```

# Box plots of nutritional content by food category across combined years (2008-2020)

```{r}
distinct_food_categ =
  final_cal_df %>% 
  drop_na(rest_type) %>% 
   mutate(
    food_category = 
      recode(food_category, "Appetizers & sides" = "Appetizers & Sides")
    ) %>% 
  distinct(food_category)

final_cal_df %>% 
  ggplot(aes(x = food_category, y = calories, fill = rest_type)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 6)) +
  labs(
    x = "Food Category",
    y = "Calories",
    title = "Distribution of Calories By Food Category")
  
# why are there NAs 
```

# counts
```{r}
final_cal_df %>%
  group_by(rest_type) %>%
  count()
```
# Most Common Restaurant in New York City 

```{r}
rest_freq =
  rest_locations %>% 
  group_by(restaurant) %>% 
  mutate(restaurant = recode(restaurant, "APPLEBEES" = "APPLEBEE'S", "AUNTIE ANNES" = "AUNTIE ANNE'S PRETZELS", "Pizza Hut" = "PIZZA HUT", "POPEYE'S" = "POPEYES", "WingStop" = "WINGSTOP"))

distinct_rest_freq =
  rest_freq %>% 
  distinct(restaurant)
  

rest_freq %>% 
  ggplot(aes(x = restaurant)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 6)) +
  labs(
    x = "Restaurant",
    y = "Frequency",
    title = "Most Common Restaurant Across New York City")
  
## put this in plotly with fill as zipcode and restaurant type...?
```

# 54 restaurants in nyc from rest_inspec dataframe --> are they all in menustat?
```{r}
final_cal_df %>% 
  filter(restaurant == "MCDONALD'S" | "POPEYES" | "KFC" | "BOSTON MARKET" | "CHIPOTLE MEXICAN GRILL" | "WHITE CASTLE" | "TIM HORTONS" | "LITTLE CAESARS" | "WINGSTOP" | "TACO BELL" | "SUBWAY" | "WENDY'S" | "SHAKE SHACK" | "PAPA JOHN'S" 
IHOP
16
PIZZA HUT
17
APPLEBEE'S
18
PANDA EXPRESS
19
CHECKERS
20
ARBY'S
21
BURGER KING
22
STARBUCKS COFFEE
23
PANERA BREAD
24
TROPICAL SMOOTHIE CAFE
25
KRISPY KREME
26
FIVE GUYS FAMOUS BURGERS AND FRIES
27
BUFFALO WILD WINGS
28
DOMINOS
29
OUTBACK STEAKHOUSE
30
MOE'S SOUTHWEST GRILL
31
AUNTIE ANNE'S PRETZELS
32
CHICK-FIL-A
33
SBARRO
34
DUNKIN’
35
HARDEE
36
HOOTERS
37
TGI FRIDAY'S
38
JAMBA JUICE
39
MARCO'S
40
RED LOBSTER
41
THE CAPITAL GRILLE
42
DAVE & BUSTER'S
43
CHUCK E. CHEESE'S
44
PERKINS RESTAURANT & BAKERY
45
JIMMY JOHNS
46
BASKIN ROBBINS
47
P.F.CHANG'S
48
DICKEY'S BARBEQUE PIT
49
THE CHEESECAKE FACTORY
50
CALIFORNIA PIZZA KITCHEN
51
DAIRY QUEEN GRILL & CHILL
52
QDOBA
53
SONIC
54
LONGHORN STEAKHOUSE)
```

