---
title: "exploratory_analysis"
author: 
date: "2022-12-01"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)
library(ggplot2)
library(patchwork)
library(dplyr)
library(readxl)
library(tidycensus)
library(rvest)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
```

```{r load datasets}
menustat_df = read_csv("data/menustat_combined.csv")
rest_locations = read_csv("data/restaurant_location.csv") 
```

# Box plot of nutritional content by restaurant type across combined years (2008-2020)

```{r}
avg_cal =
  menustat_df %>% 
  group_by(rest_type) ##why am i getting an NA for restaurant type?

avg_cal %>% 
  ggplot(aes(x = rest_type, y = calories)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Calories Per Food Item",
    title = "Distribution of Calories in Food Items Served by Restaurant Type") 


## filter by nas just to see what is coming up as that for restaurant type 
rest_type_na =
  menustat_df %>% 
  filter(
    rest_type == ""
  ) ## nothing...confused 
```
## what items are over 5500 calories each?

```{r}
cal_dense = 
  avg_cal %>% 
  mutate(
    as.numeric(calories)
  ) %>% 
  filter(
    calories > 5500
  )

fast_cal_dense = 
  cal_dense %>% 
  filter(
    rest_type == "fast food"
  ) %>% 
  select(rest_type, item_name, calories)

fast_cal_dense_plot =
  fast_cal_dense %>% 
  ggplot(aes(x = item_name, y = calories)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 6)) +
  labs(
    x = "Item Name",
    y = "Calories per Item",
    title = "Food Items > 5500 calories in Fast Food Restaurants"
  ) 

fast_cal_dense_plot

sit_cal_dense = 
  cal_dense %>% 
  filter(
    rest_type == "sit-down"
  ) %>% 
  select(rest_type, item_name, calories)
  
sit_cal_dense_plot =
sit_cal_dense %>% 
  ggplot(aes(x = item_name, y = calories)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 6)) +
  labs(
    x = "Item Name",
    y = "Calories per Item",
    title = "Food Items > 5500 calories in Sit-Down Restaurants"
  ) 

sit_cal_dense_plot

(fast_cal_dense_plot + sit_cal_dense_plot)
  
cal_dense %>% 
  ggplot(aes(x = item_name, y = calories, color = rest_type)) +
  geom_point(alpha = .5) +
  theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 6)) +
  labs(
    x = "Item Name",
    y = "Calories per Item",
    title = "Scatterplot of Food Items > 5500 calories "
  ) 

## again i think these would read better if interactive 
```

## let's look at if distribution of calories changed over the years (2008, 2012, 2016, 2020)

### 2008
```{r}
cal_2008 =
  avg_cal %>% 
  filter(
    year == "2008"
  )

cal_08_plot =
  cal_2008 %>% 
  ggplot(aes(x = rest_type, y = calories)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Calories Per Food Item",
    title = "Distribution of Calories in Food Items Served by Restaurant Type in 2008") 

cal_08_plot

cal_2012 =
  avg_cal %>% 
  filter(
    year == "2012"
  )

cal_12_plot =
  cal_2012 %>% 
  ggplot(aes(x = rest_type, y = calories)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Calories Per Food Item",
    title = "Distribution of Calories in Food Items Served by Restaurant Type in 2012") 

cal_12_plot

cal_2016 =
  avg_cal %>% 
  filter(
    year == "2016"
  )

cal_16_plot =
  cal_2016 %>% 
  ggplot(aes(x = rest_type, y = calories)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Calories Per Food Item",
    title = "Distribution of Calories in Food Items Served by Restaurant Type in 2016") 

cal_16_plot

cal_2020 =
  avg_cal %>% 
  filter(
    year == "2020"
  )

cal_20_plot =
  cal_2020 %>% 
  ggplot(aes(x = rest_type, y = calories)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Calories Per Food Item",
    title = "Distribution of Calories in Food Items Served by Restaurant Type in 2020") 

cal_20_plot

(cal_08_plot +  cal_12_plot + cal_16_plot + cal_20_plot)

## use patchwork to put all the years side by side 
## i want this to be interactive so that people can hover and see what item is that many calories 
## add another link on your personal website to mess around with the flex dashboard 
```
In 2008 see that the max calorie item recorded was almost 3000 cals. 
In 2016 see the NAs come back....huh... 

## boxplot of changing distribution of fat content over the years 

```{r}
fat_2008 = 
  menustat_df %>% 
  mutate(
    as.numeric(trans_fat)
  ) %>% 
  filter(
    year == "2008")

  
fat_08_plot =
  fat_2008 %>% 
  ggplot(aes(x = rest_type, y = trans_fat)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Trans Fat Per Food Item",
    title = "Distribution of Trans Fat in Food Items Served by Restaurant Type in 2008") 

fat_08_plot

fat_2012 = 
  menustat_df %>% 
  mutate(
    as.numeric(trans_fat)
  ) %>% 
  filter(
    year == "2012")

  
fat_12_plot =
  fat_2012 %>% 
  ggplot(aes(x = rest_type, y = trans_fat)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Trans Fat Per Food Item",
    title = "Distribution of Trans Fat in Food Items Served by Restaurant Type in 2012") 

fat_12_plot


fat_2016 = 
  menustat_df %>% 
  mutate(
    as.numeric(trans_fat)
  ) %>% 
  filter(
    year == "2016")

  
fat_16_plot =
  fat_2016 %>% 
  ggplot(aes(x = rest_type, y = trans_fat)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Trans Fat Per Food Item",
    title = "Distribution of Trans Fat in Food Items Served by Restaurant Type in 2016") 

fat_16_plot


fat_2020 = 
  menustat_df %>% 
  mutate(
    as.numeric(trans_fat)
  ) %>% 
  filter(
    year == "2020")

  
fat_20_plot =
  fat_2020 %>% 
  ggplot(aes(x = rest_type, y = trans_fat)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Trans Fat Per Food Item",
    title = "Distribution of Trans Fat in Food Items Served by Restaurant Type in 2020") 

fat_20_plot


(fat_08_plot +  fat_12_plot + fat_16_plot + fat_20_plot)
```
### boxplot of protein content over the years by restaurant type 

```{r}
prot_2008 = 
  menustat_df %>% 
  mutate(
    as.numeric(protein)
  ) %>% 
  filter(
    year == "2008")

  
prot_08_plot =
  prot_2008 %>% 
  ggplot(aes(x = rest_type, y = protein)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Protein Per Food Item",
    title = "Distribution of Protein in Food Items Served by Restaurant Type in 2008") 

prot_08_plot

prot_2012 = 
  menustat_df %>% 
  mutate(
    as.numeric(protein)
  ) %>% 
  filter(
    year == "2012")

  
prot_12_plot =
  prot_2012 %>% 
  ggplot(aes(x = rest_type, y = protein)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Protein Per Food Item",
    title = "Distribution of Protein in Food Items Served by Restaurant Type in 2012") 

prot_12_plot


prot_2016 = 
  menustat_df %>% 
  mutate(
    as.numeric(protein)
  ) %>% 
  filter(
    year == "2016")

  
prot_16_plot =
  prot_2016 %>% 
  ggplot(aes(x = rest_type, y = protein)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Protein Per Food Item",
    title = "Distribution of Protein in Food Items Served by Restaurant Type in 2016") 

prot_16_plot


prot_2020 = 
  menustat_df %>% 
  mutate(
    as.numeric(protein)
  ) %>% 
  filter(
    year == "2020")

  
prot_20_plot =
  prot_2020 %>% 
  ggplot(aes(x = rest_type, y = protein)) +
  geom_boxplot() +
  labs(
    x = "Restaurant Type",
    y = "Protein Per Food Item",
    title = "Distribution of Protein in Food Items Served by Restaurant Type in 2020") 

prot_20_plot


(prot_08_plot +  prot_12_plot + prot_16_plot + prot_20_plot)
```

# Box plots of nutritional content by food category across combined years (2008-2020)

```{r}
distinct_food_categ =
  menustat_df %>% 
  distinct(food_category)

menustat_df %>% 
  ggplot(aes(x = food_category, y = calories, fill = rest_type)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1, size = 6)) +
  labs(
    x = "Food Category",
    y = "Calories",
    title = "Distribution of Calories By Food Category")
  
# why are there NAs 
```

# counts
```{r}
menustat_df %>%
  group_by(rest_type) %>%
  count()
```

