---
title: "Proposal: Evaluating Nutritional Content of Top National Food Chains in New York City (2008-2020)"
author: ""
date: "2022-11-10"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Group Members

* Raenita Spriggs (rs4390)
* Kevin Patterson (kpp2126)
* Lindsey Covell (ltc2121)
* Gabriela Jackson (glij2117)
* Alana Ferris (af3341)

## Motivation for the Project

The growing number of fast-food chains in the United States has pushed the US into an era of food "convenience." Since the industry’s inception in 1927, there are approximately 150 thousand convenience stores across the US as of 2020.<sup>[1](https://www.emerald.com/insight/content/doi/10.1108/EUM0000000002953/full/html),[2](https://www.statista.com/topics/3869/convenience-stores-in-the-us/),[3](https://www.convenience.org/Research/FactSheets/IndustryStoreCount)</sup> Their presence is ubiquitous; however, little is known about the general distribution of fast-food chains in the US. Spatial access to convenience stores from schools in low and high-income neighborhoods, and dietary quality have been explored. <sup>[4](https://escholarship.org/uc/item/93f0g6z8),[5](https://link.springer.com/article/10.1007/s12061-013-9095-6),[6](https://www.frontiersin.org/articles/10.3389/fpubh.2022.857754/full)</sup> In cities like New York, there seems to be a new chain on every corner of the city. They provide quick and cheap food options, alongside fast food chains, which could be the only option for many communities. 

Convenience stores and others alike are known to sell food products high in fats, sugars, and salts.<sup>[7](https://pubmed.ncbi.nlm.nih.gov/22898159/)</sup> The presence of convenience stores is associated with higher BMI, higher obesity, and increased hypertension.<sup>[8](https://pubmed.ncbi.nlm.nih.gov/16530621/),[9](https://pubmed.ncbi.nlm.nih.gov/17884578/)</sup> Residents of color living in rural, low-income neighborhoods have inadequate access to supermarkets, chain grocery stores, and diversity in food products.<sup>[10](https://www.ajpmonline.org/article/S0749-3797(08)00838-6/fulltext)</sup> We are motivated by these findings to investigate trends in NYC. 

Our project seeks to investigate at the zip-code level in New York, NY:

1. The general spatial distribution of convenience stores in relation to demographic composition;
2. Trends of nutritional content of food served across years (2008-2020)

## Intended Final Products
- A report containing the introduction, methods, results, conclusions, and future directions
- A dashboard showcasing maps for NYC at the zip code level

## Anticipated data sources
- [NYC DOHMH MenuStat nutritional database](https://www.menustat.org/data.html) 
- [NYC DOHMH Restaurant Inspections database](https://data.cityofnewyork.us/Health/Restaurants-rolled-up-/59dk-tdhz/data)
- [NY U.S. census mean income data](https://data.census.gov/cedsci/table?q=population%20demographics%20income&g=0400000US36%248600000&tid=ACSST5Y2020.S1902) 
- [NY U.S. census demographic and housing estimate data](https://data.census.gov/cedsci/table?q=population%20demographics%20&g=0400000US36%248600000&tid=ACSDP5Y2020.DP05)

## Planned analysis
- OLS regressions at the zip code level estimating food quality index based on ~%race, total population, average income, %convenient store, %trans fat/total fat content across all food outlets for a zip code

## Visualizations
- Box plot of nutritional content by food category and restaurant type across combined years (2008-2020)
- Spaghetti plot of nutritional content by restaurant and coded by restaurant type
- Bar plots of top 15 NYC restaurant chains
- Maps stratified by restaurant type illustrating restaurants with the most/least nutrient-dense menu items according to a national standard diet for the average adult (CDC MyPlate).

## Coding Challenges
- Significant missingness for 10 nutritional variables (`Total_fat`, `Saturated_fat`, `Trans_fat`, `Cholesterol`, `Sodium`, `Potassium`, `Carbohydrates`, `Protein`, `Sugar`, `Dietary_fiber`); could be addressed using conditional mean imputation depending on differences in missingness based on food category.
- An idea of the number of NA’s in 2018 alone:
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
df_2018 = readxl::read_excel("data/annual_data2018.xlsx") %>%
  janitor::clean_names()

df_2018 %>%
	summarise(across(everything(), ~sum(is.na(.x)))) %>%
  select(-c(menu_item_id, year, restaurant, food_category, item_name, item_description, kids_meal, limited_time_offer, regional, shareable)) %>%
	knitr::kable()
```
- Create factor vector that indicates restaurant type: fast food, convenience store, sit-down restaurant

## Timeline
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
tibble(Date = c("Nov 7-10", "Nov 10", "Nov 11", "Nov 14", "Nov 15 - 17", "Nov 15/16", "Nov 18", "Dec 5", "Dec 6", "Dec 10", "Dec 16"),
       Task = c("Brainstorm", 
                "Finish draft proposal", 
                "Submit proposal",
                "Designate tasks",
                "Cleaning Data",
                "Project Review Meeting",
                "Start Analysis",
                "Finalize code",
                "Begin and upload website",
                "Report, webpage, screencast, peer assessment",
                "In class discussion"),
       Due = c("-", 
                "-", 
                "Nov 12 by 1:00 pm",
                "End of day",
                "Nov 15-17",
                "Nov 15 or 16",
                "-",
                "-",
                "End of day",
                "Dec 10 by 11:59 pm",
                "Dec 15")) %>%
  knitr::kable()
```

