---
title: "Exploratory Analysis Flexdashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
---

```{r setup, include=FALSE}
library(flexdashboard)
library(plotly)
library(tidyverse)
menustat_df = read_csv("data/menustat_combined.csv")
rest_locations = read_csv("data/restaurant_location.csv") 

```
## when i try to knit this document it isn't working, using the code `rmarkdown::render("exploratory_anlalysis_flexdashboard.Rmd", output_format = "flexdashboard::flex_dashboard")` which i got from class website 

Initial Visualizations
======================================================================

Column {data-width=650}
-----------------------------------------------------------------------

### Distribution of Total Calories By Restaurant Type

```{r}
menustat_df %>% 
  group_by(rest_type) %>% 
  plot_ly(x = ~rest_type, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

Column {data-width=350}
-----------------------------------------------------------------------

### Items With Calorie Counts in Top 5% of Restaurants 

```{r}
menustat_df %>% 
  group_by(rest_type) %>% 
  filter(quantile(calories, 0.95) < calories) %>% 
  plot_ly(x = ~rest_type, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Frequency of Restaurants Across NYC

```{r}
rest_freq %>% 
  count(restaurant) %>% 
  mutate(neighbourhood = fct_reorder(restaurant, n)) %>% 
  plot_ly(x = ~restaurant, y = ~n, color = ~restaurant, type = "bar", colors = "viridis")
```


Calorie Distribution By Restaurant Type
======================================================================

Column {data-width=650}
-----------------------------------------------------------------------


### Calorie Distribution in 2008

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2008") %>% 
  plot_ly(x = ~rest_type, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calorie Distribution in 2012

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2012") %>% 
  plot_ly(x = ~rest_type, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calorie Distribution in 2016

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2016") %>% 
  plot_ly(x = ~rest_type, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calorie Distribution in 2020

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2020") %>% 
  plot_ly(x = ~rest_type, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```


Fat Distribution By Restaurant Type
======================================================================

Column {data-width=650}
-----------------------------------------------------------------------


### Fat Distribution in 2008

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2008") %>% 
  plot_ly(x = ~rest_type, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Fat Distribution in 2012

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2012") %>% 
  plot_ly(x = ~rest_type, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Fat Distribution in 2016

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2016") %>% 
  plot_ly(x = ~rest_type, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Fat Distribution in 2020

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2020") %>% 
  plot_ly(x = ~rest_type, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```


Protein Distribution By Restaurant Type
======================================================================

Column {data-width=650}
-----------------------------------------------------------------------


### Protein Distribution in 2008

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2008") %>% 
  plot_ly(x = ~rest_type, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in 2012

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2012") %>% 
  plot_ly(x = ~rest_type, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in 2016

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2016") %>% 
  plot_ly(x = ~rest_type, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in 2020

```{r}
final_cal_df %>% 
  group_by(rest_type) %>% 
  filter(year == "2020") %>% 
  plot_ly(x = ~rest_type, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```


Nutritional Content by Food Category
======================================================================

Column {data-width=650}
-----------------------------------------------------------------------

### Calories Distribution by Food Category Overlayed

```{r}
final_cal_df %>% 
  drop_na(rest_type) %>% 
  plot_ly(x = ~food_category, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calories Distribution in Beverages and Desserts Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "beverages_dessert") %>% 
  plot_ly(x = ~food_category, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calories Distribution in Convenience Stores

```{r}
final_cal_df %>% 
  filter(rest_type == "convenience") %>% 
  plot_ly(x = ~food_category, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calories Distribution in Fast Food Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "fast food") %>% 
  plot_ly(x = ~food_category, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Calories Distribution in Sit-down Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "sit-down") %>% 
  plot_ly(x = ~food_category, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Fat Distribution by Food Category Overlayed

```{r}
final_cal_df %>% 
  drop_na(rest_type) %>% 
  plot_ly(x = ~food_category, y = ~calories, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Fat Distribution in Beverages and Desserts Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "beverages_dessert") %>% 
  plot_ly(x = ~food_category, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Fat Distribution in Convenience Stores

```{r}
final_cal_df %>% 
  filter(rest_type == "convenience") %>% 
  plot_ly(x = ~food_category, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Fat Distribution in Fast Food Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "fast food") %>% 
  plot_ly(x = ~food_category, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Fat Distribution in Sit-down Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "sit-down") %>% 
  plot_ly(x = ~food_category, y = ~trans_fat, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution by Food Category Overlayed

```{r}
final_cal_df %>% 
  drop_na(rest_type) %>% 
  plot_ly(x = ~food_category, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in Beverages and Desserts Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "beverages_dessert") %>% 
  plot_ly(x = ~food_category, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in Convenience Stores

```{r}
final_cal_df %>% 
  filter(rest_type == "convenience") %>% 
  plot_ly(x = ~food_category, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in Fast Food Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "fast food") %>% 
  plot_ly(x = ~food_category, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```

### Protein Distribution in Sit-down Restaurants

```{r}
final_cal_df %>% 
  filter(rest_type == "sit-down") %>% 
  plot_ly(x = ~food_category, y = ~protein, color = ~rest_type, text = ~item_name, type = "box", colors = "viridis")
```
